Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data

### 2015-01-27 Paper/ms.tex
Occam's razor and nu Oph numbers

The unknown parameters are:
\begin{eqnarray}
\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, m_0,
\sigma_{\rm extra}, \nu\right\}
\end{eqnarray}
where $N$ is the number of planets,
$\boldsymbol{\alpha} = \{\mu_P, \sigma_P, \mu_A\}$ are hyperparameters
hyperparameters used to define the prior for the properties of the planets,
and $\psi_i = \{P_i, A_i, e_i, \phi_i, \omega_i\}$
are the properties of planet $i$.
The parameter $m_0$ describes a DC offset in the data, and
$\sigma_{\rm extra}$ and $\nu$ are parameters of the noise distribution
which are discussed further below.

@@

The posterior distribution for the logarithms of the periods is shown in
Figure~\ref{fig:nu_oph_periods}. As in Section~\ref{sec:fake_data},
the posterior samples for all periods were combined to make this figure, which
shows several prominent peaks. 
The two peaks with vertical dashed lines are the commonly accepted periods of
530 and 3190 days, and the other prominent peaks (i.e. signals which have a
moderate probability of existence) have periods of 36.11 $\pm$ 0.034 days,
75.58 $\pm$ 0.80 days, and 1709 $\pm$ 183 days.
The probability that
at least one of these ``extra'' signals exists, given by the fraction of
samples that have one of these signals, is 85\%.
Given that they exist, their amplitudes are low, around 5-40 metres per second,
which we note is above the noise level. An example model fit to the data is
shown in Figure~\ref{fig:nuoph}.

@@

The posterior distribution for $N$ is shown in Figure~\ref{fig:gliese581_N},
and shows strong evidence for at least eight periods
($P(N \leq 7 | D) = 0.012$). Some authors \citep[e.g.][]{2011A&A...528L...5T}
recommend that the probability of $N$ planets should be $\geq$ 150 times
greater than the probability of $N-1$ planets existing before making a
claim that $N$ planets have been definitively detected. Such a decision rule
is presumably equivalent to a utility function where false positives are much
worse than false negatives. We note that applying this rule to our results, we
would assert $N=6$, even though this has a tiny probability.

It is now recognised that
there are many possible sources for oscillations in a data set and not all
such oscillations should be claimed as planets \citep{2014Sci...345..440R}. 
Our model cannot distinguish between oscillations due to planets and
oscillations due to stellar activity: any oscillations found in the dataset
will be described as ``planets'' by the model. Another important consideration
is the physical stability of the orbital system which is ignored by this type
of analysis \citep{orbital}. However it is interesting that
we find many more signals in the data than previous authors.
By inspecting the posterior distribution for the periods
(Figure~\ref{fig:gliese581_periods}), we see that only four of the periods
are well determined, corresponding to the known periods of Gliese 581 b, c, d,
and e. The other ``periods'' are very uncertain. It is quite likely that a
non-sinusoidal signal due to stellar activity is being modelled as several
periods \citep{astero}, and if the model were extended to include a ``stochastic''
oscillation \citep{gaussproc}, the number of periods detected may be reduced
substantially. An example model fit for Gliese 581 is shown in
Figure~\ref{fig:gliese581}.

@@

>>>>

The unknown parameters are:
\begin{eqnarray}
\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, m_0,
\sigma_{\rm extra}, \nu\right\}
\end{eqnarray}
where $N$ is the number of planets,
$\boldsymbol{\alpha} = \{\mu_P, \sigma_P, \mu_A\}$ are hyperparameters
hyperparameters used to define the prior for the properties of the planets,
and $\psi_i = \{P_i, A_i, e_i, \phi_i, \omega_i\}$
are the properties of planet $i$.
The parameter $m_0$ describes a DC offset in the data, and
$\sigma_{\rm extra}$ and $\nu$ are parameters of the noise distribution
which are discussed further below. Note that our parameter $\omega_i$ is
standard, however $\phi_i$ is non-standard because we assert that $\phi_i=0$
always implies the signal is at its maximum at $t=0$. Our parameter space is
equivalent to the standard one, we are just using a different coordinate system.

@@

The posterior distribution for the logarithms of the periods is shown in
Figure~\ref{fig:nu_oph_periods}. As in Section~\ref{sec:fake_data},
the posterior samples for all periods were combined to make this figure, which
shows several prominent peaks. 
The two peaks with vertical dashed lines are the commonly accepted periods of
530 and 3190 days, and the other prominent peaks (i.e. signals which have a
moderate probability of existence) have periods of 36.11 $\pm$ 0.034 days,
75.58 $\pm$ 0.80 days, and 1709 $\pm$ 183 days. As with any MCMC output,
if we are interested in the probability of any proposition $Q$ (for example,
``$Q \equiv$ a planet exists with period between 35 and 37 days''), we can
calculate the proportion of the posterior samples for which $Q$ is true, which
(if we have a lot of samples) is a Monte Carlo estimate of the posterior
probability of $Q$. For $\nu$ Oph, we calculated the probability that
at least one of these ``extra'' signals (beyond the two commonly accepted ones)
exists, as 85\%.
Given that they exist, their amplitudes are low, around 5-40 metres per second,
which we note is above the noise level. An example model fit to the data is
shown in Figure~\ref{fig:nuoph}.

@@

The posterior distribution for $N$ is shown in Figure~\ref{fig:gliese581_N},
and shows strong evidence for at least eight periods
($P(N \leq 7 | D) = 0.012$). Some authors \citep[e.g.][]{2011A&A...528L...5T}
recommend that the probability of $N$ planets should be $\geq$ 150 times
greater than the probability of $N-1$ planets existing before making a
claim that $N$ planets have been definitively detected. Such a decision rule
is presumably equivalent to a utility function where false positives are much
worse than false negatives. We note that applying this rule to our results, we
would assert $N=6$, even though this has a very small posterior probability.

It is now recognised that
there are many possible sources for oscillations in a data set and not all
such oscillations should be claimed as planets. 
Our model cannot distinguish between oscillations due to planets and
oscillations due to stellar activity: any oscillations found in the dataset
will be described as ``planets'' by the model. Another important consideration
is the physical stability of the orbital system which is ignored by this type
of analysis \citep{orbital}. However it is interesting that
we find many more signals in the data than previous authors.
By inspecting the posterior distribution for the periods
(Figure~\ref{fig:gliese581_periods}), we see that only four of the periods
are well determined, corresponding to the known periods of Gliese 581 b, c, d,
and e. The other ``periods'' are very uncertain. It is quite likely that a
non-sinusoidal signal due to stellar activity is being modelled as several
periods \citep{astero}, and if the model were extended to include a ``stochastic''
oscillation \citep{gaussproc}, the number of periods detected may be reduced
substantially. With these kinds of models, the posterior distribution for $N$
can also be influenced by the prior for the amplitudes $A_i$. Many authors
assign independent broad priors to the amplitudes, and this causes the
``Occam's razor'' penalty for adding extra signals to be quite strong. Since we
use a hierarchical prior for the amplitudes, if some amplitudes are found to
be low, $\mu_A$ will become small. When $\mu_A$ is small, it is likely that any
extra signals will have small amplitudes, so the ``Occam's razor'' effect is
weaker. An example model fit for Gliese 581 is shown in
Figure~\ref{fig:gliese581}.

@@

As with $\nu$ Oph, we can calculate the posterior probability of any hypothesis
about Gliese 581 by computing the fraction of the posterior samples that have
that property. The posterior probabilities for planets b, c, d, and e, are
close to 1. The posterior probability a signal exists with $\log_{10}(P)
\in [1.55, 1.57]$ is 88\%, and the probability for a signal with
$\log_{10}(P) \in [2.6, 2.8]$ is 85\%.
